---
title: "Clothing Purchasing and Disposal Habits of ETH Zurich Members"
date: today
author:
  - name: Anahita Darvish
    orcid: 0000-0003-2196-5015
    email: adarvish@student.ethz.ch
    affiliation: 
      - name: ETH Zurich
        url: https://ethz.ch/de.html
format: html
execute:
  warning: false
editor_options: 
  chunk_output_type: console
#bibliography: references.bib
editor: visual
---

# Introduction

4-5 sentences

Circular fashion is becoming increasingly important as fast fasion industries and incorrect clothing disposal methods are damaging the environment. This report aims to analyze the clothing purchasing and disposal habits of ETH Zurich members and make suggestions on how best promote circular fashion at ETH.

# Methods

3-5 sentences

Data was collected through a survey questionnaire created on Google forms with questions relating to demographics, environmental awareness, clothing purchasing habits, and clothing disposal habits. Responses were collected in the following two ways: 1) through a link shared on WhatsApp to ETH students and 2) through a QR code given out to any ETH members on Polyterrace. A total of 20 responses were recorded and the subsequent raw data was exported in Google sheets. Data cleaning and analysis was done in Quarto using R scripts.

# Results

Each figure/table has 2-3 sentences of interpretation

Summary table with statistics

```{r}
#| label: tbl-stats
#| tbl-cap: Summary Statistics of Key Survey Variables

library(tidyverse)
library(gt)
library(here)

data <- read_csv(here("data/processed", "clothing_data_cleaned.csv"))

# Compute summary statistics
summary_table <- data |> 
  summarise(
    Mean = c(mean(new_items, na.rm = TRUE),
             mean(env_conscious, na.rm = TRUE),
             mean(second_hand, na.rm = TRUE)),
    
    Median = c(median(new_items, na.rm = TRUE),
               median(env_conscious, na.rm = TRUE),
               median(second_hand, na.rm = TRUE)),
    
    SD = c(sd(new_items, na.rm = TRUE),
           sd(env_conscious, na.rm = TRUE),
           sd(second_hand, na.rm = TRUE)),
    
    Min = c(min(new_items, na.rm = TRUE),
            min(env_conscious, na.rm = TRUE),
            min(second_hand, na.rm = TRUE)),
    
    Max = c(max(new_items, na.rm = TRUE),
            max(env_conscious, na.rm = TRUE),
            max(second_hand, na.rm = TRUE))
  ) |> 
  mutate(Variable = c("New Items Purchased", "Environmental Consciousness", "Percent Second-Hand Purchases")) |> 
  select(Variable, everything())

# Display using gt
summary_table |> 
  gt() |> 
  fmt_number(columns = where(is.numeric), decimals = 2)


```

```         
```

Data visualization of purchasing habits by income level

```{r}
#| label: fig-income
#| fig-cap: Purchasing Habits by Income Level

library(tidyverse)
library(ggplot2)
library(ggthemes)

# Reshape purchasing method columns
purchasing_long <- data |> 
  pivot_longer(cols = starts_with("get_"),
               names_to = "method",
               values_to = "response") |> 
  filter(response == "yes")

# Clean method labels
purchasing_long$method <- purchasing_long$method |> 
  str_replace("get_", "") |> 
  str_replace_all("_", " ") |> 
  str_to_title()

# Plot
ggplot(data = purchasing_long, 
       mapping = aes(x = income,
                     fill = method)) +
  geom_bar(position = "fill")+
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Income (CHF/month)",
       y = "Proportion of Respondents",
       fill = "Acquisition Method")+
  theme_minimal() +
  scale_fill_colorblind()
```

Data visualization for disposal habits by gender

```{r}
#| label: fig-continent
#| fig-cap: Disposal Habits by Continent

# Reshape disposal method columns to long format
disposal_long <- data |> 
  pivot_longer(cols = starts_with("disposal_"),
               names_to = "method",
               values_to = "response") |> 
  filter(response == "yes")

# Clean method labels
disposal_long$method <- disposal_long$method |> 
  str_replace("disposal_", "") |> 
  str_replace_all("_", " ") |> 
  str_to_title()

# Count total number of respondents per continent
respondents_per_continent <- data |> 
  group_by(continent) |> 
  summarise(n_total = n(), .groups = "drop")

# Count yes responses per disposal method per continent
lollipop_data <- disposal_long |> 
  group_by(continent, method) |> 
  summarise(n_yes = n(), .groups = "drop") |> 
  left_join(respondents_per_continent, by = "continent") |> 
  mutate(percent = (n_yes / n_total) * 100)

# Plot lollipop chart
ggplot(lollipop_data, aes(x = percent, y = method, color = continent)) +
  geom_point(size = 4) +
  geom_segment(aes(x = 0, xend = percent, y = method, yend = method)) +
  labs(
    x = "Percent of Respondents (per Continent)",
    y = "Disposal Method",
    color = "Continent"
  ) +
  theme_minimal() +
  scale_color_colorblind()

```

```{r}
# Create data frame with each disposal method and initialize percentage
disposal_habits_df <- data.frame(
  disposal_method = c("Donate", "Flea Market", "Family & Friends",
                      "Sell Online", "Recyling Bins", "Regular Trash", 
                      "Third Party Textile Recycling"),
  disposal_percentages = numeric(7)
)

# Calculate percentage of respondent's disposal habits and assign to correct place in data frame
disposal_habits_df[1, "disposal_percentages"] <- sum(data$disposal_thrift == "yes") / count(data) * 100
disposal_habits_df[2, "disposal_percentages"] <- sum(data$disposal_flea_market == "yes") / count(data) * 100
disposal_habits_df[3, "disposal_percentages"] <- sum(data$disposal_fam_friends == "yes") / count(data) * 100
disposal_habits_df[4, "disposal_percentages"] <- sum(data$disposal_sell_online == "yes") / count(data) * 100
disposal_habits_df[5, "disposal_percentages"] <- sum(data$disposal_recycling_bins == "yes") / count(data) * 100
disposal_habits_df[6, "disposal_percentages"] <- sum(data$disposal_trash == "yes") / count(data) * 100
disposal_habits_df[7, "disposal_percentages"] <- sum(data$disposal_textile_center == "yes") / count(data) * 100

```

```{r}
# Create data frame with each purchasing method and initialize percentage
purchasing_habits_df <- data.frame(
  purchasing_method = c("Thrift", "Hand-me-downs", "Clothing Swaps",
                      "Gifts", "Make Clothing"),
  purchasing_percentages = numeric(5)
)

# Calculate percentage of respondent's purchasing habits and assign to correct place in data frame
purchasing_habits_df[1, "purchasing_percentages"] <- sum(data$get_thrift == "yes") / count(data) * 100
purchasing_habits_df[2, "purchasing_percentages"] <- sum(data$get_hand_me_down == "yes") / count(data) * 100
purchasing_habits_df[3, "purchasing_percentages"] <- sum(data$get_swaps == "yes") / count(data) * 100
purchasing_habits_df[4, "purchasing_percentages"] <- sum(data$get_gifts == "yes") / count(data) * 100
purchasing_habits_df[5, "purchasing_percentages"] <- sum(data$get_make_clothes == "yes") / count(data) * 100
```

```{r}
# Calculate how many respondents purchase new clothes 
brand_sust_filtered <- data |> 
  filter(!is.na(brand_sust))

new_clothes_df <- data.frame(
  buyers_percentage = numeric(1),
  brand_sust_percentage = numeric(1))

new_clothes_df[1, "buyers_percentage"] <- sum(data$new_clothes == "yes") / count(data) * 100
new_clothes_df[1, "brand_sust_percentage"] <-  sum(brand_sust_filtered$brand_sust == "yes") / count(brand_sust_filtered) * 100
```

```{r}
# Mean percentage of second hand clothing purchases
second_hand_purchases_percentage <- mean(data$second_hand)

data |> 
  count(motivation)

data |> 
  count(disposal)

data |> 
  count(disposal_reason)
```

# Conclusion

Bulleted format

```{r}

```
