---
title: "Clothing Purchasing and Disposal Habits of ETH Zurich Members"
date: today
author:
  - name: Anahita Darvish
    orcid: 0000-0003-2196-5015
    email: adarvish@student.ethz.ch
    affiliation: 
      - name: ETH Zurich
        url: https://ethz.ch/de.html
format: html
execute:
  warning: false
editor_options: 
  chunk_output_type: console
#bibliography: references.bib
editor: visual
---

# Introduction

4-5 sentences

Circular fashion is becoming increasingly important as fast fashion industries and incorrect clothing disposal methods are damaging the environment. This report aims to analyze the clothing purchasing and disposal habits of ETH Zurich members and make suggestions on how best promote circular fashion at ETH.

# Methods

3-5 sentences

Data was collected using a survey questionnaire created in Google Forms, which included questions about demographics, environmental awareness, clothing purchasing habits, and clothing disposal practices. Responses were gathered in two ways: (1) via a link shared on WhatsApp with ETH students, and (2) through a QR code distributed to ETH members on Polyterrace. A total of 20 responses were recorded, and the raw data was exported to Google Sheets. Data cleaning and analysis were conducted in Quarto using R scripts.

# Results

Each figure/table has 2-3 sentences of interpretation

## Study limitations

Convenience sampling and size of response pool may affect the generalizability of results. Proportions of respondents in each income bracket and continent of origin were not controlled.

Summary table with statistics

@tbl-stats displays key summary statistics of the survey respondents. On average, respondent's purchased 6.4 new clothing items in the last semester, respondents also demonstrated a relatively high value for environmental consciousness with an average value of 3.3 on a scale from 0-5 in which 0 = not environmentall conscious at all and 5 = extrememly environmentally conscious. The average respondent purchased 36.90% second-hand clothing in the past semester, however there was a very high standard deviation in which some responder's purchased 0% second-hand clothing and other purchased 87.5%.

```{r}
#| label: tbl-stats
#| tbl-cap: Summary Statistics of Key Survey Variables

library(tidyverse)
library(gt)
library(here)

data <- read_csv(here("data/processed", "clothing_data_cleaned.csv"))

# Compute summary statistics
summary_table <- data |> 
  summarise(
    Mean = c(mean(new_items, na.rm = TRUE),
             mean(env_conscious, na.rm = TRUE),
             mean(second_hand, na.rm = TRUE)),
    
    Median = c(median(new_items, na.rm = TRUE),
               median(env_conscious, na.rm = TRUE),
               median(second_hand, na.rm = TRUE)),
    
    SD = c(sd(new_items, na.rm = TRUE),
           sd(env_conscious, na.rm = TRUE),
           sd(second_hand, na.rm = TRUE)),
    
    Min = c(min(new_items, na.rm = TRUE),
            min(env_conscious, na.rm = TRUE),
            min(second_hand, na.rm = TRUE)),
    
    Max = c(max(new_items, na.rm = TRUE),
            max(env_conscious, na.rm = TRUE),
            max(second_hand, na.rm = TRUE))
  ) |> 
  mutate(Variable = c("New Items Purchased", "Environmental Consciousness", "Percent Second-Hand Purchases")) |> 
  select(Variable, everything())

# Display using gt
summary_table |> 
  gt() |> 
  fmt_number(columns = where(is.numeric), decimals = 2)


```

Data visualization of motivations and reasoning

```{r}
#| label: fig-motivation
#| fig-cap: Motivations and Disposal Reasons in Clothing Consumption

library(tidyverse)
library(patchwork)  # for combining plots

# Count motivations
motivation_plot <- data %>%
  count(motivation) %>%
  ggplot(aes(x = fct_reorder(motivation, n), y = n)) +
  geom_col(fill = "#56B4E9") +
  coord_flip() +
  labs(title = "Motivations for Purchasing Second-Hand",
       x = NULL, y = "Number of Respondents") +
  theme_minimal()

# Count disposal reasons
disposal_reason_plot <- data %>%
  count(disposal_reason) %>%
  ggplot(aes(x = fct_reorder(disposal_reason, n), y = n)) +
  geom_col(fill = "#E69F00") +
  coord_flip() +
  labs(title = "Reasons for Disposing of Clothing",
       x = NULL, y = "Number of Respondents") +
  theme_minimal()

# Combine them stacked
motivation_plot / disposal_reason_plot

```

Data visualization of purchasing habits by income level

The effects of income level on purchasing habits are explored in @fig-income. The results demonstrated that thrift and vintage stores are a primary purchasing method amount all the income groups explored. It was also observed that respondent's with over 3000 CHF monthly income were they only income group who did not partake in clothing swaps and were less reliant on hand-me-downs as a clothing aquistion method. This could suggest that high income individuals

```{r}
#| label: fig-income
#| fig-cap: Purchasing Habits by Income Level

library(tidyverse)
library(ggplot2)
library(ggthemes)

# Reshape purchasing method columns
purchasing_long <- data |> 
  pivot_longer(cols = starts_with("get_"),
               names_to = "method",
               values_to = "response") |> 
  filter(response == "yes")

# Clean method labels
purchasing_long$method <- purchasing_long$method |> 
  str_replace("get_", "") |> 
  str_replace_all("_", " ") |> 
  str_to_title()

# Plot
ggplot(data = purchasing_long, 
       mapping = aes(x = income,
                     fill = method)) +
  geom_bar(position = "fill")+
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Income (CHF/month)",
       y = "Proportion of Respondents",
       fill = "Acquisition Method")+
  theme_minimal() +
  scale_fill_colorblind()
```

Data visualization for disposal habits by continent

@fig-continent displays clothing disposal methods separated by the respondent's continent of origin. Based on the results, North American respondent's did not use clothing recycling bins or textile recycling centers as a means for disposal, which may suggest a lack of education about the important of textile recycling. Whereas respondents from Asia were least likely to throw clothes away in the regular trash, suggesting they may have a higher education on the subject due to the continent's booming textile industry.

```{r}
#| label: fig-continent
#| fig-cap: Disposal Habits by Continent

# Reshape disposal method columns to long format
disposal_long <- data |> 
  pivot_longer(cols = starts_with("disposal_"),
               names_to = "method",
               values_to = "response") |> 
  filter(response == "yes")

# Clean method labels
disposal_long$method <- disposal_long$method |> 
  str_replace("disposal_", "") |> 
  str_replace_all("_", " ") |> 
  str_to_title()

# Count total number of respondents per continent
respondents_per_continent <- data |> 
  group_by(continent) |> 
  summarise(n_total = n(), .groups = "drop")

# Count yes responses per disposal method per continent
lollipop_data <- disposal_long |> 
  group_by(continent, method) |> 
  summarise(n_yes = n(), .groups = "drop") |> 
  left_join(respondents_per_continent, by = "continent") |> 
  mutate(percent = (n_yes / n_total) * 100)

# Plot lollipop chart
ggplot(lollipop_data, aes(x = percent, y = method, color = continent)) +
  geom_point(size = 4) +
  geom_segment(aes(x = 0, xend = percent, y = method, yend = method)) +
  labs(
    x = "Percent of Respondents (per Continent)",
    y = "Disposal Method",
    color = "Continent"
  ) +
  theme_minimal() +
  scale_color_colorblind()

```


```{r}
# Create data frame with each disposal method and initialize percentage
disposal_habits_df <- data.frame(
  disposal_method = c("Donate", "Flea Market", "Family & Friends",
                      "Sell Online", "Recyling Bins", "Regular Trash", 
                      "Third Party Textile Recycling"),
  disposal_percentages = numeric(7)
)

# Calculate percentage of respondent's disposal habits and assign to correct place in data frame
disposal_habits_df[1, "disposal_percentages"] <- sum(data$disposal_thrift == "yes") / count(data) * 100
disposal_habits_df[2, "disposal_percentages"] <- sum(data$disposal_flea_market == "yes") / count(data) * 100
disposal_habits_df[3, "disposal_percentages"] <- sum(data$disposal_fam_friends == "yes") / count(data) * 100
disposal_habits_df[4, "disposal_percentages"] <- sum(data$disposal_sell_online == "yes") / count(data) * 100
disposal_habits_df[5, "disposal_percentages"] <- sum(data$disposal_recycling_bins == "yes") / count(data) * 100
disposal_habits_df[6, "disposal_percentages"] <- sum(data$disposal_trash == "yes") / count(data) * 100
disposal_habits_df[7, "disposal_percentages"] <- sum(data$disposal_textile_center == "yes") / count(data) * 100

```

```{r}
# Create data frame with each purchasing method and initialize percentage
purchasing_habits_df <- data.frame(
  purchasing_method = c("Thrift", "Hand-me-downs", "Clothing Swaps",
                      "Gifts", "Make Clothing"),
  purchasing_percentages = numeric(5)
)

# Calculate percentage of respondent's purchasing habits and assign to correct place in data frame
purchasing_habits_df[1, "purchasing_percentages"] <- sum(data$get_thrift == "yes") / count(data) * 100
purchasing_habits_df[2, "purchasing_percentages"] <- sum(data$get_hand_me_down == "yes") / count(data) * 100
purchasing_habits_df[3, "purchasing_percentages"] <- sum(data$get_swaps == "yes") / count(data) * 100
purchasing_habits_df[4, "purchasing_percentages"] <- sum(data$get_gifts == "yes") / count(data) * 100
purchasing_habits_df[5, "purchasing_percentages"] <- sum(data$get_make_clothes == "yes") / count(data) * 100
```

```{r}
# Calculate how many respondents purchase new clothes 
brand_sust_filtered <- data |> 
  filter(!is.na(brand_sust))

new_clothes_df <- data.frame(
  buyers_percentage = numeric(1),
  brand_sust_percentage = numeric(1))

new_clothes_df[1, "buyers_percentage"] <- sum(data$new_clothes == "yes") / count(data) * 100
new_clothes_df[1, "brand_sust_percentage"] <-  sum(brand_sust_filtered$brand_sust == "yes") / count(brand_sust_filtered) * 100
```


# Conclusions

-   Second-hand shopping is becoming increasingly popular amoung all income groups

-   Higher income individuals are less likely to participate in clothing swaps and recieve hand-me-downs

-   North Americans lack education over proper clothing disposal habits

-   
