---
title: "Clothing Consumption and Disposal Habits of ETH Zurich Members"
date: today
author:
  - name: Anahita Darvish
    orcid: 0009-0000-4405-9889
    email: adarvish@student.ethz.ch
    affiliation: 
      - name: ETH Zurich
        url: https://ethz.ch/de.html
format: html
toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
bibliography: references.bib
editor: visual
---

# Introduction

Textile waste has become an enormous burden on the environment as fast fashion industries have risen in popularity. The low-price, trendy clothing offered by fast fashion industries, including ZARA and H&M, are produced with lower quality materials leading to shorter clothing lifetimes and larger amounts of textile waste. Circular fashion, including the reuse and recycling of clothing, is the key to combating this issue. This report aims to answer the following research question:

What are the clothing disposal habits of ETH Zurich students, and how are they influenced by their clothing consumption patterns (new vs second-hand) during the past semester?

By analyzing the clothing purchasing and disposal behaviors of ETH Zurich members, recommendations for the promotion of a circular fashion mindset at ETH can be made.

# Methods

Data was collected using a survey questionnaire created in Google Forms, which included questions about demographics (age, gender, education, income, continent of origin), environmental awareness, clothing purchasing habits, and clothing disposal practices. All clothing purchasing and disposal questions were dichotomous choice questions (yes or no). Responses were gathered in two ways: (1) via a link shared on WhatsApp with ETH students, and (2) through a QR code distributed to ETH members on Polyterrace over a one-hour period. A total of 20 responses were recorded, and the raw data was exported to Google Sheets. Data cleaning and analysis were conducted in Quarto using R scripts.

# Results

@tbl-stats displays key summary statistics from the survey. On average, respondents purchased 6.4 new clothing items in the last semester and demonstrated a relatively high level of environmental consciousness, with an average score of 3.3 on a 0–5 scale (0 = not at all environmentally conscious, 5 = extremely conscious). The average share of second-hand clothing purchases was 36.9%, though this varied widely — some respondents reported 0%, while others reported as high as 87.5%, suggesting significant diversity in purchasing behavior.

```{r}
#| label: tbl-stats
#| tbl-cap: Summary Statistics of Key Survey Variables

library(tidyverse)
library(gt)
library(here)

data <- read_csv(here("data/processed", "clothing_data_cleaned.csv"))

# Compute summary statistics
summary_table <- data |> 
  summarise(
    Mean = c(mean(new_items, na.rm = TRUE),
             mean(env_conscious, na.rm = TRUE),
             mean(second_hand, na.rm = TRUE)),
    
    Median = c(median(new_items, na.rm = TRUE),
               median(env_conscious, na.rm = TRUE),
               median(second_hand, na.rm = TRUE)),
    
    SD = c(sd(new_items, na.rm = TRUE),
           sd(env_conscious, na.rm = TRUE),
           sd(second_hand, na.rm = TRUE)),
    
    Min = c(min(new_items, na.rm = TRUE),
            min(env_conscious, na.rm = TRUE),
            min(second_hand, na.rm = TRUE)),
    
    Max = c(max(new_items, na.rm = TRUE),
            max(env_conscious, na.rm = TRUE),
            max(second_hand, na.rm = TRUE))
  ) |> 
  mutate(Variable = c("New Items Purchased", "Environmental Consciousness", "Percent Second-Hand Purchases")) |> 
  select(Variable, everything())

# Display using gt
summary_table |> 
  gt() |> 
  fmt_number(columns = where(is.numeric), decimals = 2)


```

@fig-motivation highlights the most common motivations behind second-hand clothing purchases, as well as the primary reasons for clothing disposal among respondent's, the majority of whom are young adults. Interestingly, most participants reported choosing second-hand clothing for its style and uniqueness, rather than for economic or environmental reasons. Implying that among ETH members, vintage or second-hand clothing is seen as fashionable, rather than simply practical. This aligns with research showing that shifting fashion norms are helping drive the popularity of second-hand clothing among younger, educated populations [@ferraro2016]. On the disposal side, most respondents indicated they discard clothing only when it is damaged or worn out, reflecting a more sustainable attitude in which garments are kept until the end of their useful life, rather than discarded due to changing trends.

```{r}
#| label: fig-motivation
#| fig-cap: Motivations and Disposal Reasons in Clothing Consumption

library(patchwork)  # for combining plots

# Count motivations
motivation_plot <- data |> 
  count(motivation, age) |> 
  ggplot(mapping = aes(x = fct_reorder(motivation, n), 
                       y = n,
                       fill = age)) +
  geom_col() +
  coord_flip() +
  labs(title = "Motivations for Purchasing Second-Hand",
       x = NULL, 
       y = "Number of Respondents") +
  theme_minimal()

# Count disposal reasons
disposal_reason_plot <- data |> 
  count(disposal_reason, age) |> 
  ggplot(mapping = aes(x = fct_reorder(disposal_reason, n), 
                       y = n,
                       fill = age)) +
  geom_col() +
  coord_flip() +
  labs(title = "Reasons for Disposing of Clothing",
       x = NULL, 
       y = "Number of Respondents") +
  theme_minimal()

# Combine them stacked
motivation_plot / disposal_reason_plot

```

@fig-income explores the effect of income level on clothing consumption practices. Thrift and vintage stores are widely used across all income groups, highlighting the growing normalization of second-hand fashion. In @markova2018 the authors describe how thrifting has evolved into "a lifestyle for consumers of the urban city", reflecting a cultural shift that extends beyond economic necessity. However, respondents with a monthly income above 3000 CHF were the only group that did not participate in clothing swaps, and they showed less reliance on hand-me-downs compared to lower-income groups. This may suggest that higher income members of ETH, such as PhD students or faculty, are less engaged with circular fashion practices due to reduced exposure — for example, they may be less likely to see swap flyers on campus or hear about events through peer networks.

```{r}
#| label: fig-income
#| fig-cap: Purchasing Habits by Income Level

library(ggplot2)
library(ggthemes)

# Reshape purchasing method columns
purchasing_long <- data |> 
  pivot_longer(cols = starts_with("get_"),
               names_to = "method",
               values_to = "response") |> 
  filter(response == "yes")

# Clean method labels
purchasing_long$method <- purchasing_long$method |> 
  str_replace("get_", "") |> 
  str_replace_all("_", " ") |> 
  str_to_title()

# Plot
ggplot(data = purchasing_long, 
       mapping = aes(x = income,
                     fill = method)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Income (CHF/month)",
       y = "Proportion of Respondents",
       fill = "Acquisition Method") +
  theme_minimal() +
  scale_fill_colorblind()
```

@fig-continent shows disposal habits by respondents’ continent of origin. North American respondents did not report using clothing recycling bins or textile recycling centers, which may suggest lower awareness of textile recycling infrastructure. In contrast, respondents from Asia were least likely to dispose of clothing via regular trash, potentially indicating greater awareness of textile waste or cultural practices influenced by the region’s prominent role in the global textile industry. This aligns with global findings, where China stands out for its progress in circular economy efforts, while the USA struggles to translate awareness into action [@gonella2024].

```{r}
#| label: fig-continent
#| fig-cap: Disposal Habits by Continent

# Reshape disposal method columns to long format
disposal_long <- data |> 
  pivot_longer(cols = starts_with("disposal_"),
               names_to = "method",
               values_to = "response") |> 
  filter(response == "yes")

# Clean method labels
disposal_long$method <- disposal_long$method |> 
  str_replace("disposal_", "") |> 
  str_replace_all("_", " ") |> 
  str_to_title()

# Count total number of respondents per continent
respondents_per_continent <- data |> 
  group_by(continent) |> 
  summarise(n_total = n(), .groups = "drop")

# Count yes responses per disposal method per continent
lollipop_data <- disposal_long |> 
  group_by(continent, method) |> 
  summarise(n_yes = n(), .groups = "drop") |> 
  left_join(respondents_per_continent, by = "continent") |> 
  mutate(percent = (n_yes / n_total) * 100)

# Plot lollipop chart
ggplot(lollipop_data, aes(x = percent, y = method, color = continent)) +
  geom_point(size = 4) +
  geom_segment(aes(x = 0, xend = percent, y = method, yend = method)) +
  labs(
    x = "Percent of Respondents (per Continent)",
    y = "Disposal Method",
    color = "Continent"
  ) +
  theme_minimal() +
  scale_color_colorblind()

```

# Conclusions

-   The average ETH Zurich member demonstrates a moderatly high level of environmental consciousness and purchases over a third of their clothing from second-hand sources; however, large variablity exists in the percentage of second-hand clothing purchased.

-   Young adults at ETH believe second-hand shopping to be trendy, as their key motivation for shopping second-hand is the style and uniqueness.

-   The main reason to dispose of clothing is due to damage or end of life of the garment, suggesting a sustainable mindset in the ETH community.

-   Thrifting is becoming increasingly popular among all income groups, supporting the normalization of second-hand shopping in urban areas.

-   Higher income individuals are less likely to participate in circular fashion practices on campus due to lack of exposure to such events.

-   Respondents of North American origin demonstrate the least participation in circular fashion practices likely due to a lack of information or awareness.

-   Respondents of Asian origin show the greatest participation in circular fashion practices and are the least likely to dispose of clothing in the regular trash.

Recommendations:

-   Flyers given to all new members of ETH Zurich with information regarding best clothing disposal practices.

-   Integration of circular fashion into campus culture by hosting joint events with different student organizations or departments.

-   Create an ETH Zurich webpage that lists second-hand stores, swap events, disposal guides, and facts about clothing waste.

Study limitations:

Convenience sampling and size of response pool may affect the generalizability of results. Proportions of respondents in each age group, income bracket and continent of origin were not controlled.

# References
